---
title: "Seascape Stats"
author: "Lachlan Phillips"
date: '2022-06-29'
output:
  rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
library(sjPlot)
library(nlme)
# Note: rmdformats is also a dependency

height=3
```


Data is a data frame off all CTD casts with environmental and acoustic summary statistics. 

```{r}
cast.df = readRDS('../data/processed/cast_df_stats.rds')
```

The goals of these GLM models are:

- Show the relationship between environment and aggregation characteristics
- Show the differing relationships between environment and aggregation characteristics between the two seascape classifications.

I am showing model results in tables by using the `tab_model` function from the library **sjPlot**. The below is a wrapper function. Feel free to modify it if you think other stats from the GLMs should be displayed.
```{r}
# wrapper for showing model results
tab_model.wrap <- function(fit){
  tab_model(fit, show.aic = T, show.se=T, show.ci=F)
}
```


# Cluster GLMs

--- 

Again not really sure how to to interpret these. But my take was that there is only significant differences between clusters for density and biomass. Is this your take? Matches the plots. 

## **depth ~ temp x cluster**
```{r}
fit <- glm(swarm_mean_depth ~ temp_mean*cluster, data=cast.df, family=Gamma(link='log'))
tab_model.wrap(fit)
anova(fit)

ggplot(cast.df, aes(temp_mean, swarm_mean_depth, col = cluster)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = Gamma(link = "log")))
```

<br><br>

## **coverage ~ temp x cluster**
```{r}
## use Box-Cox transformation for area 
bc <- MASS::boxcox(area_corrected_area_sum ~ temp_mean*cluster, data = cast.df, plotit = FALSE)
## lambda is 0.2222 for these data - close to 1/3 but residuals are a bit closer to Normal
lambda <- bc$x[which.max(bc$y)] 

## sjPlot::tab_model() doesn't like transformations in model calls, so modify the data.frame
cast.df <- dplyr::mutate(cast.df, 
                         transf_area_sum = ((area_corrected_area_sum^lambda - 1)/lambda))

fit <- glm(transf_area_sum ~ temp_mean*cluster, data=cast.df, family = "gaussian")
tab_model.wrap(fit)

ggplot(cast.df, aes(temp_mean, transf_area_sum, col = cluster)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

<br><br>

## **density ~ temp x cluster**
```{r}
fit <- glm(Sv_mean_mean ~ temp_mean*cluster, data=cast.df, family='gaussian')
tab_model.wrap(fit)

ggplot(cast.df, aes(temp_mean, Sv_mean_mean, col = cluster)) + geom_point() + geom_smooth(method = "glm")
```


<br><br><br>


# Temperature and Salinity

--- 

## **temp ~ salt**
```{r}
fit <- glm(temp_mean ~ salt_mean, data=cast.df, family='gaussian')
tab_model.wrap(fit)

ggplot(cast.df, aes(salt_mean, temp_mean)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family='gaussian'))
```


<br><br><br>


## **temp ~ salt x survey**
```{r}
fit <- glm(temp_mean ~ salt_mean, data = cast.df, family = gaussian)
tab_model.wrap(fit)
summary(fit)

ggplot(cast.df, aes(salt_mean, temp_mean, col = survey)) + 
  geom_point() + 
  geom_smooth(method = "glm")
```

<br><br><br>

# Crossshore Gradient

---

## **temp ~ crossShoreDistance**
```{r}
df.split <- split(cast.df, cast.df$survey)
for (i in 1:length(df.split)){
  message('\n', names(df.split)[i])
  fit <- glm(temp_mean ~ crossShoreDistance, data=df.split[[i]])
  #print(tab_model.wrap(fit))
  print(summary(fit))
}

ggplot(cast.df, aes(crossShoreDistance, temp_mean)) + 
  geom_point() + 
  geom_smooth(method = "glm") +
  facet_wrap('survey')
```

<br><br>

## **coverage ~ crossShoreDistance**
```{r}
df.split <- split(cast.df, cast.df$survey)
for (i in 1:length(df.split)){
  message('\n', names(df.split)[i])
  fit <- glm(transf_area_sum ~ crossShoreDistance, data=df.split[[i]])
  #print(tab_model.wrap(fit))
  print(summary(fit))
}

ggplot(cast.df, aes(crossShoreDistance, transf_area_sum)) + 
  geom_point() + 
  geom_smooth(method = "glm") +
  facet_wrap('survey')
```

<br><br>

## **density ~ crossShoreDistance**
```{r}
df.split <- split(cast.df, cast.df$survey)
for (i in 1:length(df.split)){
  message('\n', names(df.split)[i])
  fit <- glm(Sv_mean_mean ~ crossShoreDistance, data=df.split[[i]])
  #print(tab_model.wrap(fit))
  print(summary(fit))
}

ggplot(cast.df, aes(crossShoreDistance, Sv_mean_mean)) + 
  geom_point() + 
  geom_smooth(method = "glm") +
  facet_wrap('survey')
```